---
title: "ML project"
author: "Jing Huo"
date: "January 27, 2016"
output: html_document
---

The goal of this project is to predict the manner in which they did the exercise. 

Here are the libraries used.
```{r}
library(kernlab)
library(caret)
library(randomForest)
```

# Loading data

```{r}
traindata <- read.csv("C:\\Users\\JHuo\\Desktop\\coursera\\machine Learning\\pml-training.csv", as.is=TRUE)
testdata <- read.csv("C:\\Users\\JHuo\\Desktop\\coursera\\machine Learning\\pml-testing.csv", as.is=TRUE)
```

# PreProcessing
The first 7 columns are identifier columns to be removed. Also, empty columns and those with lots of NAs are removed by visual observation.

```{r}

traindata <- traindata[, c(8:11, 37:49, 60:68, 84:86, 102,113:124, 140, 151:ncol(traindata) )]
testdata <- testdata[, c(8:11, 37:49, 60:68, 84:86, 102,113:124, 140, 151:ncol(testdata) )]
```

Remove those columns that have zero variance.
```{r}
nsv <- nearZeroVar(traindata)
traindata_nsv <- traindata
if(length(nsv) > 0) traindata_nsv <- traindata[, -nsv]
```

Split the traindata into training and validation sets
```{r}
inTrain <- createDataPartition(y=traindata_nsv$classe, p=0.7, list=FALSE)
training <- traindata[inTrain,]
testing <- traindata[-inTrain,]
```

Remove those rows with NAs
```{r}
cc <- complete.cases(training)
training <- training[cc,]
```

Rescale into N(0,1)
```{r}
preObj <- preProcess(training[,-53], method=c("center","scale"))
train_scale <- predict(preObj, training[,-53])
test_scale <- predict(preObj, testing[,-53])
```

Oberserve the dataset and see the correlations between variables.
```{r}
M <- abs(cor(train_scale))
diag(M)<-0
which(M>0.8, arr.ind=T)

plot(train_scale$accel_belt_x, train_scale$pitch_belt)
```

# Building the model

Here I apply a random forest classifier, with 200 trees and 5 variables for each tree.

```{r}
clf_200 <- randomForest(factor(training$classe)~., data=train_scale, ntree=200, nodesize=5)
```

# Results

Here is the list of variable importance.

```{r}
varImpPlot(clf_200, sort=TRUE, main="variable importance from random forest")
```

Here are the results from validation set.
```{r}
resProb <- predict(clf_200, test_scale, type="response")
xtab=table(resProb, testing$classe)
print(confusionMatrix(xtab))
```


